How Large a Sample Size is Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates?

### 2023-03-09 Manuscript/manuscript.tex
Few Manuscript Chagnes

As described in the last section, the block bootstrap method can be used to
estimate a parameter $\theta$ of an AR(1) process. This method can be
simulated with the tsboot function from the R boot package \citep{boot}
\citep{davison}. In this study, $\theta$ is composed of the target parameters
$\mu = 0$, $\sigma_{x}$, and $\phi$. The sample mean $\bar{x}$ can be used as
the statistic to estimate $\mu$, the sample standard deviation $S_x$ can be
used as the statistic to estimate $\sigma$, and the autocorrelation function
of the sample at lag 1 $\rho_1$ can be used as the statistic to estimate
$\phi$. In the R implementation, these three statistics are incorporated into
a function that will apply each statistic to the original sample and to each
bootstrapped pseudo-sample. From a sample of length $n$, two types of block
bootstrap 95\% confidence intervals can be constructed for each target
parameter:
$CI_{\hat{\theta}_{n}}$ and $CI_{BCA, \hat{\theta}_{n}}$.

Each of these 6 intervals either recovers $\theta$ or does not. Remember that
95\% of many of a certain type of 95\% confidence intervals constructed from
samples of the same $n$ should recover $\theta$. To see if this is
approximately the case, one can run 1000 replications of the block bootstrap
procedure, and for each interval-$\theta$ pair, the rate at which that type of
interval recovers $\theta$ can be recorded. The coverage rate is, of course, a
point estimate of a proportion, so a 95\% confidence interval of this coverage
proportion (with $n_{CI} = 1000$) can be constructed, and if the proportion
.95 is included in the interval, the block bootstrap method is likely working
well. If all values in the interval are below .95, the results suggest that the
method either is providing an inaccurate estimation, is underestimating the
process' variability, or a combination of both. If all values in the interval
are above .95, the results suggest that the method is overestimating the
process' variability.

@@

The figures below were created using the R ggplot2 package \citep{ggplot2}.
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_n.4}
  \caption{At $\phi = -.4$, both intervals covered $\mu$ correctly as low as
    $n = 100$. $CI_{\hat{\theta}_{n}}$ seems to cover $\sigma$ well as low as
    $n = 400$. $CI_{BCA, \hat{\theta}_{n}}$ under-covers $\sigma$, but there
    is no clear pattern of deterioration of coverage, so one can expect that
    increasing $n$ will eventually result in proper coverage of $\sigma$.
    $CI_{\hat{\theta}_{n}}$ covers $\phi$ well with no deterioration, and this
    is true as low as $n = 200$. $CI_{BCA, \hat{\theta}_{n}}$ greatly
    under-covers $\phi$, and there is a clear pattern of deterioration of
    coverage.}
  \label{fig:plot_n.4}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_n.2}
  \caption{At $\phi = -.2$, both intervals covered $\mu$ correctly as low as
    $n = 100$. Both intervals seem to cover $\sigma$ generally well, although
    $CI_{BCA, \hat{\theta}_{n}}$ has somewhat lower coverage rates.
    $CI_{\hat{\theta}_{n}}$ covers $\phi$ well with no deterioration of
    coverage, and this is true as low as $n = 200$.
    $CI_{BCA, \hat{\theta}_{n}}$ greatly under-covers $\phi$, and there is a
    clear coverage deterioration trend as $n$ increases.}
  \label{fig:plot_n.2}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_0}
  \caption{At $\phi = 0$, both intervals covered $\mu$ correctly as low as
    $n = 500$. $CI_{\hat{\theta}_{n}}$ seems to cover $\sigma$ well as low as 
    $n = 400$. $CI_{BCA, \hat{\theta}_{n}}$ slightly under-covers $\sigma$, but
    there is no clear pattern of deterioration of coverage, so one can expect
    that increasing $n$ will eventually result in proper coverage of $\sigma$.
    $CI_{\hat{\theta}_{n}}$ seems to cover $\phi$ generally well at $n = 400$,
    but $CI_{BCA, \hat{\theta}_{n}}$ under-covers $\phi$, although the
    method's performance seems to improve as $n$ increases.}
  \label{fig:plot_0}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.2}
  \caption{At $\phi = .2$, both intervals covered $\mu$ correctly as low as
    $n = 300$. Both intervals seem to cover $\sigma$ generally well, although
    $CI_{BCA, \hat{\theta}_{n}}$ has somewhat lower coverage rates.
    $CI_{\hat{\theta}_{n}}$ covers $\phi$ well with no deterioration of
    coverage, and this is true as low as $n = 200$.
    $CI_{BCA, \hat{\theta}_{n}}$ greatly under-covers $\phi$, but there is no
    clear coverage deterioration trend as $n$ increases.}
  \label{fig:plot_.2}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.4}
  \caption{At $\phi = .4$, both intervals seem to cover $\mu$ generally well
    as low as $n = 400$. $CI_{\hat{\theta}_{n}}$ covers $\sigma$ well as low
    as $n = 400$. $CI_{BCA, \hat{\theta}_{n}}$ under-covers $\sigma$, but
    there is no clear sign of coverage deterioration, so one can expect that
    increasing $n$ will eventually result in proper coverage of $\sigma$.
    $CI_{\hat{\theta}_{n}}$ covers $\phi$ well with no deterioration of
    coverage, and this is true as low as $n = 200$.
    $CI_{BCA, \hat{\theta}_{n}}$ greatly under-covers $\phi$ with a clear
    coverage deterioration trend as $n$ increases.}
  \label{fig:plot_.4}
\end{figure}

As shown in the above figures, only $CI_{\hat{\theta}_{n}}$ managed to solve
the deterioration problem with respect to recovering $\phi$, indicating that
this interval should be used as opposed to the more complex
$CI_{BCA, \hat{\theta}_{n}}$ when estimating $\phi$. Increasing $n$ will not
fix the performance of $CI_{BCA, \hat{\theta}_{n}}$ because the assumption that
the coverage rate will approach an upper limit of .95 as $n$ increases is
invalid. As expected for $CI_{\hat{\theta}_{n}}$, a higher $\phi$ demands a
larger $n$ to estimate all parameters, but a negative $\phi$ may require a
lesser $n$. For example, for both $\phi = 0.2$ and $0.4$, $n = 100$ was
sufficient to recover $\mu$ at a rate of approximately 95\%.

@@

The motivation for the study is the idea that block bootstrap procedure will
perfectly estimate a parameter of a time series given an infinitely large
sample. The goal for this study was to find what is a good enough finite
sample length for block bootstrap estimation to recover the parameter of a
time series at an acceptable rate. Again, this relies on the assumption that
there is a size $n$ large enough for the method to work: that is, the method's
performance improves as $n$ increases. Out of the two types of intervals used
in this study, this assumption was found to hold true with respect to
estimating $\phi$ only for $CI_{\hat{\theta}_{n}}$, whereas
$CI_{BCA, \hat{\theta}_{n}}$ exhibited coverage deterioration as $n$
increased. When using $CI_{\hat{\theta}_{n}}$ and $\phi$ is unknown, the
results of this study suggest that an $n$ of around 500 may be necessary to
estimate $\mu$, and an $n$ of around 400 may be necessary to estimate
$\sigma$. If $\phi$ is already known, a lesser $n$ may be adequate to estimate
these parameters. When $\phi$ was negative, a $n$ of around 100 was sufficient
to estimate $\mu$. However, in real world applications, $\phi$ is more
commonly found to be positive, so a larger $n$ is more likely to be necessary.
Lastly, to estimate $\phi$, an $n$ of around 400 may be necessary. This
information could prove to be useful for researching using block bootstrap
estimation of time series in domains such as econometrics. Future studies
could investigate if there are types of block bootstrap intervals
constructions that would more appropriately recover the parameters of a time
series. One could also investigate the $n$ needed to make inferences about
other forms of serially dependent data such as an MA process.

>>>>

As described in the last section, the block bootstrap method can be used to
estimate a parameter $\theta$ of an AR(1) process. This method can be
simulated with the tsboot function from the R boot package \citep{boot}
\citep{davison1997bootstrap}. In this study, $\theta$ is composed of the target parameters
$\mu = 0$, $\sigma_{x}$, and $\phi$. The sample mean $\bar{x}$ can be used as
the statistic to estimate $\mu$, the sample standard deviation $S_x$ can be
used as the statistic to estimate $\sigma$, and the autocorrelation function
of the sample at lag 1 $\rho_1$ can be used as the statistic to estimate
$\phi$. In the R implementation, these three statistics are incorporated into
a function that will apply each statistic to the original sample and to each
bootstrapped pseudo-sample. From a sample of length $n$, three types of block
bootstrap 95\% confidence intervals can be constructed for each target
parameter:
Percentile, $\hat{\theta}_{n}$-Centered,
and $BC_a$.

Each of these 9 intervals either recovers $\theta$ or does not. Remember that
95\% of many of a certain type of 95\% confidence intervals constructed from
samples of the same $n$ should recover $\theta$. To see if this is
approximately the case, one can run 1000 replications of the block bootstrap
procedure, and for each interval-$\theta$ pair, the rate at which that type of
interval recovers $\theta$ can be recorded. The coverage rate is, of course, a
point estimate of a proportion, so a 95\% confidence interval of this coverage
proportion (with $n_{CI} = 1000$) can be constructed, and if the proportion
.95 is included in the interval, the block bootstrap method is likely working
well. If all values in the interval are below .95, the results suggest that the
method either is providing an inaccurate estimation, is underestimating the
process' variability, or a combination of both. If all values in the interval
are above .95, the results suggest that the method is overestimating the
process' variability.

@@

The figures below which are grouped by the target $\theta$  were created using
the R ggplot2 package \citep{ggplot2}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_mu}
  \caption{At $\phi = -.4, -.2$, all intervals seem to cover $\mu$ correctly
    as low as $n = 100$. At $\phi = 0$, an $n$ of 500 seems large enough to
    cover $\mu$ acceptably regardless of interval type. At $\phi = .2$,
    $n = 300$ seems to be sufficient to appropriately recover $\mu$ for all
    intervals. A much larger $n$ seems to be necessary to recover $\mu$ at a
    $\phi$ as high as $.4$.}
  \label{fig:plot_mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_sigma}
  \caption{At $\phi = -.4$, the percentile and the
    $\hat{\theta}_{n}$-centered intervals recovered $\mu$ well as low as
    $n = 400$, whereas the $BC_a$ interval undercovered $\mu$ for all $n$
    observed. At $\phi = -.2$, the percentile and the
    $\hat{\theta}_{n}$-centered intervals recovered $\mu$ well as low as
    $n = 500$, but the $BC_a$ coverage rates were lower.}
  \label{fig:plot_sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_phi}
  \caption{}
  \label{fig:plot_phi}
\end{figure}

As shown in the above figures, only $\hat{\theta}_{n}$-centered intervals
managed to solve the deterioration problem with respect to recovering $\phi$,
indicating that this interval should be used as opposed to the uncentered
percentile interval or the more complex $BC_a$ interval when estimating
$\phi$. Increasing $n$ will not fix the performance of either of these
intervals because the assumption that the coverage rate will approach an upper
limit of .95 as $n$ increases is invalid. As expected for
$\hat{\theta}_{n}$-centered intervals, a higher positive $\phi$ demands a
larger $n$ to estimate all parameters, but a negative $\phi$ may require a
lesser $n$. For example, for both $\phi = 0.2$ and $0.4$, $n = 100$ was
sufficient to recover $\mu$ at a rate of approximately 95\%.

@@

The motivation for the study is the idea that block bootstrap procedure will
perfectly estimate a parameter of a time series given an infinitely large
sample. The goal for this study was to find what is a good enough finite
sample length for block bootstrap estimation to recover the parameter of a
time series at an acceptable rate. Again, this relies on the assumption that
there is a size $n$ large enough for the method to work: that is, the method's
performance improves as $n$ increases. Out of the two types of intervals used
in this study, this assumption was found to hold true with respect to
estimating $\phi$ only for $\hat{\theta}_{n}$-centered intervals, whereas both
percentile $BC_a$ intervals exhibited coverage deterioration as $n$
increased. When using $\hat{\theta}_{n}$-centered intervals and $\phi$ is
unknown, the results of this study suggest that an $n$ of around 500 may be
necessary to estimate $\mu$, and an $n$ of around 400 may be necessary to
estimate $\sigma$. If $\phi$ is already known, a lesser $n$ may be adequate to
estimate these parameters. When $\phi$ was negative, a $n$ of around 100 was
sufficient to estimate $\mu$. However, in real world applications, $\phi$ is
more commonly found to be positive, so a larger $n$ is more likely to be
necessary. Lastly, to estimate $\phi$, an $n$ of around 400 may be necessary.
This information could prove to be useful for researching using block bootstrap
estimation of time series in domains such as econometrics. Future studies
could investigate if there are types of block bootstrap intervals
constructions that would more appropriately recover the parameters of a time
series. One could also investigate the $n$ needed to make inferences about
other forms of serially dependent data such as an MA process.
